---
title: "01_dataset_creation"
author: "Miguel Armengol"
date: "2/20/2019"
output: html_document
---

# Libraries

```{r}
library(bigrquery)
library(summarytools)
library(readr)
library(stringr)
library(sqldf)
library(dplyr)
```


# Set up BigQuery related functions

This chunks also creates the run_query and get_sql function.

```{r setup, include=FALSE}

project_id <- "hst-953-2018"
options(httr_oauth_cache=FALSE)
run_query <- function(query){
  data <- query_exec(query, project=project_id, use_legacy_sql=FALSE,max_pages = Inf)
  return(data)
}

# function for reading sql files
getSQL <- function(filepath){
  con = file(filepath, "r")
  sql.string <- ""

  while (TRUE){
    line <- readLines(con, n = 1)

    if ( length(line) == 0 ){
      break
    }

    line <- gsub("\\t", " ", line)

    if(grepl("--",line) == TRUE){
      line <- paste(sub("--","/*",line),"*/")
    }

    sql.string <- paste(sql.string, line)
  }

  close(con)
  return(sql.string)
}
```

# Loading queries and extracting the data

Loads all queries from the sql files in the extraction folder and runs them into RBigQuey to extract the data.

```{r}
#folders contained in sql folder
#cat(list.files('sql'), sep = '\n')

# AKI related variables
baseline_creat<-run_query(getSQL("sql/aki/baseline_creat.sql"))
chronicAKI<-run_query(getSQL("sql/aki/chronicAKI.sql"))
first_rrt<-run_query(getSQL("sql/aki/first_rrt.sql"))
peakcreat48h<-run_query(getSQL("sql/aki/peakcreat48h.sql"))
peakcreat7days<-run_query(getSQL("sql/aki/peakcreat7days.sql"))

# rest of the variables
bg_24_48<-run_query(getSQL("sql/bg_24_48.sql"))
bg_first<-run_query(getSQL("sql/bg_first.sql"))
demographics<-run_query(getSQL("sql/demographics.sql"))
lab_24_48<-run_query(getSQL("sql/lab_24_48.sql"))
lab_first<-run_query(getSQL("sql/lab_first.sql"))
outcomes<-run_query(getSQL("sql/outcomes.sql"))
pasthistory<-run_query(getSQL("sql/pasthistory.sql"))
sid_24_48<-run_query(getSQL("sql/sid_24_48.sql"))
 sid_twa_24<-run_query(getSQL("sql/sid_twa_24.sql"))
temp_day1<-run_query(getSQL("sql/temp_day1.sql"))
temp_day2<-run_query(getSQL("sql/temp_day2.sql"))
urine_output_24_48<-run_query(getSQL("sql/urine_output_24_48.sql"))
vent_bin_first_2days<-run_query(getSQL("sql/vent_bin_first_2days.sql"))
vitals_day1<-run_query(getSQL("sql/vitals_day1.sql"))
vitals_day2<-run_query(getSQL("sql/vitals_day2.sql"))
```

## AKI Dataset creation

```{r}
#We need to exclude patients who already had AKI before ICU admin)
AKIdevelopment<-subset(baseline_creat, !(baseline_creat$patientunitstayid %in% chronicAKI$patientunitstayid) )
AKIdevelopment<-left_join(AKIdevelopment, peakcreat7days)
AKIdevelopment<-left_join(AKIdevelopment, peakcreat48h)
AKIdevelopment$position<-NULL
```

### Windsorization

Clean data by means of winsorization, i.e., by shrinking outlying observations to the border of the main part of the data.
The creatinine values have been winsorize replacing the extreme observations using 95% and 5% percentiles.

```{r}
AKIdevelopmentW<-AKIdevelopment
## Winsorization: Replace the extreme observations using 95% and 5% percentiles
winsorize_x = function(x, cut = 0.05){
  cut_point_top <- quantile(x, 1 - cut, na.rm = T)
  cut_point_bottom <- quantile(x, cut, na.rm = T)
  i = which(x >= cut_point_top) 
  x[i] = cut_point_top
  j = which(x <= cut_point_bottom) 
  x[j] = cut_point_bottom
  return(x)
}

AKIdevelopmentW$creat1<-winsorize_x(AKIdevelopmentW$creat1)
AKIdevelopmentW$peakcreat7d<-winsorize_x(AKIdevelopmentW$peakcreat7d)
AKIdevelopmentW$peakcreat48h<-winsorize_x(AKIdevelopmentW$peakcreat48h)

#Now we turn all emptpy values into 0

AKIdevelopmentW$creat1[is.na(AKIdevelopmentW$creat1)]<-0
AKIdevelopmentW$peakcreat7d[is.na(AKIdevelopmentW$peakcreat7d)]<-0
AKIdevelopmentW$peakcreat48h[is.na(AKIdevelopmentW$peakcreat48h)]<-0


```

###List of patients with AKI and offset

```{r}

AKIdevelopmentW['cr7undercreat1']<-AKIdevelopmentW$peakcreat7d/AKIdevelopmentW$creat1
AKIdevelopmentW['cr48hdundercreat1']<-AKIdevelopmentW$peakcreat48h-AKIdevelopmentW$creat1
AKIdevelopmentW['AKIoffset']<-NA
AKIdevelopmentW['AKIstage']<-NA

#nested if else function

i <- function(if_stat, then) {
  if_stat <- lazyeval::expr_text(if_stat)
  then    <- lazyeval::expr_text(then)
  sprintf("ifelse(%s, %s, ", if_stat, then)
}

e <- function(else_ret) {
  else_ret <- lazyeval::expr_text(else_ret)
  else_ret
}

ie <- function(...) {
  args <- list(...)
  
  for (i in 1:(length(args) - 1) ) {
      if (substr(args[[i]], 1, 6) != "ifelse") {
        stop("All but the last argument, need to be i functions.", call. = FALSE)
      }
  }
  if (substr(args[[length(args)]], 1, 6) == "ifelse"){
    stop("Last argument needs to be an e function.", call. = FALSE)
  }
  args$final <- paste(rep(')', length(args) - 1), collapse = '')
  eval_string <- do.call('paste', args)
  eval(parse(text = eval_string))
}

AKIdevelopmentW$AKIstage <- 
  ie(
    i(AKIdevelopmentW$cr7undercreat1>=3,   3),
    i(AKIdevelopmentW$cr7undercreat1>=2,   2),
    i(AKIdevelopmentW$cr7undercreat1>=1.5,   1),
    i(AKIdevelopmentW$cr48hdundercreat1>=0.3,   1),
    e(0)
  )

AKIdevelopmentW$AKIoffset <- 
  ie(
    i(AKIdevelopmentW$cr7undercreat1>=3,AKIdevelopmentW$peakcreat7d_offset),
    i(AKIdevelopmentW$cr7undercreat1>=2,AKIdevelopmentW$peakcreat7d_offset),
    i(AKIdevelopmentW$cr7undercreat1>=1.5,AKIdevelopmentW$peakcreat7d_offset),
    i(AKIdevelopmentW$cr48hdundercreat1>=0.3,   AKIdevelopmentW$peakcreat48h_offset),
    e(0)
  )

library(dplyr)
AKIlist<-AKIdevelopmentW %>%
  select(patientunitstayid,AKIstage
         # ,AKIoffset The nephrologist suggested not to use this variable since we can not infer when AKI was developed from it. Removing it to avoid confusion.
         )
```

### Final List of patients with AKI and offset (with RRT)

Offset is not reliable so we exclude it.

```{r}
# Dr Danziger said the offset is not reliable for telling when the patient exactly developed AKI so he adviced not to use it

AKIlist_final<-left_join(AKIlist,first_rrt)


AKIlist_final$AKIstage <- 
  ie(
    i(!is.na(AKIlist_final$first_rrtoffset),   3),
    e(AKIlist_final$AKIstage)
  )

# The nephrologist suggested not to use this variable since we can not infer when AKI was developed from it. Removing it to avoid confusion.
# AKIlist_final$AKIoffset <- 
#   ie(
#     i(!is.na(AKIlist_final$first_rrtoffset), AKIlist_final$first_rrtoffset),
#     e(AKIlist_final$AKIoffset)
#   )

AKIlist_final$first_rrtoffset<-NULL
AKIlist_final[is.na(AKIlist_final)]<-0
AKIlist_final$rrt_bin<-NULL
```

# Merging all datasets

We are using a left join to join them

```{r}
# left join = all.x=TRUE
sid_aki_study<-Reduce(function(...) merge(..., all.x=TRUE), list(
  demographics
, bg_first
, bg_24_48
, lab_first
, lab_24_48
, outcomes
, pasthistory
, sid_first
, sid_24_48
, sid_twa_24
, temp_day1
, temp_day2
, urine_output_24_48
, vent_bin_first_2days
, vitals_day1
, vitals_day2
, baseline_creat
, chronicAKI
, first_rrt
, peakcreat48h
, peakcreat7days
, AKIlist_final
))

cat(names(sid_aki_study),sep = '\n')
```

# Dataset report

```{r}
view(dfSummary(sid_aki_study))
```

# Export dataset

```{r}
write.csv(sid_aki_study,'sid_aki_study.csv',row.names = F)
```


# New Variables

## Data extraction from BigQuery

```{r}
admission_vars<-run_query(getSQL("sql/new_variables/admission_vars.sql"))
mean_first_day<-run_query(getSQL("sql/new_variables/mean_first_day.sql"))

```



## Merging all datasets

We are using a left join to join them

```{r}
# left join = all.x=TRUE
new_variables<-Reduce(function(...) merge(..., all.x=TRUE), list(
  admission_vars
, mean_first_day
))
```

## Dataset report

```{r}
view(dfSummary(new_variables))
```

## Export dataset

```{r}
write.csv(new_variables,'new_variables.csv',row.names = F)
```

