---
title: "Untitled"
output: html_document
---

```{r}
library(dplyr)
library(tableone)
library(psych)
library(ggplot2)
library(MASS)
library(geepack)
library(gridExtra)
library(grid)
```

```{r}
rm(list=ls())
```


# Data wrangling

```{r}
sid_aki_study<-read.csv("sid_aki_study-7.csv",header=T)
colnames(sid_aki_study)
summary(sid_aki_study$creat_first)
summary(sid_aki_study$creat_adm)
summary(sid_aki_study$creat_before_adm)

nrow(sid_aki_study)
# Admission data
df_1<-sid_aki_study%>%dplyr::select(patientunitstayid,creat_adm,creat_adm_offset,age_fixed,gender,ethnicity,weight,height,albumin_first,bilirubin_first,BUN_first,wbc_first,platelets_first,hematocrit_first,hemoglobin_first,glucose_first,unitlos_days,hosplos_days,hospitalmortality,icumortality,pasthistory_chf,pasthistory_pvd,pasthistory_hypertension,pasthistory_copd,pasthistory_diabetes,pasthistory_cld,pasthistory_malignancy,sodium_first,chloride_first,SID_first,mechvent_day01,first_rrtoffset,rrt_bin,peakcreat48h_offset,peakcreat48h,peakcreat7d,peakcreat7d_offset,peakcreat7d_to_discharge_offsetgap,AKIstage,potassium_adm,bicarbonate_adm,ua_adm,heartrate_adm,systolic_adm,temperature_adm,dialysis_in_the_icu_first7days_bin,first_dialysis_in_the_icu_first7days_offset,hospitalAdmitSource,ventdays_days,creat_first)
  
colnames(df_1)
summary(sid_aki_study$chloride_first)
summary(sid_aki_study$chloride_min_24)
summary(sid_aki_study$chloride_max_24)
summary(sid_aki_study$chloride_twa_24)


summary(sid_aki_study$albumin_first)
summary(sid_aki_study$albumin_max_24)

summary(sid_aki_study$hemoglobin_first)
summary(sid_aki_study$hemoglobin_min_24)
summary(sid_aki_study$hemoglobin_max_24)


summary(sid_aki_study$peakcreat7d)
  
#%>%mutate(aki=as.factor(aki),icumortality=as.factor(icumortality),hospitalmortality=as.factor(hospitalmortality),pasthistory_malignancy=as.factor(pasthistory_malignancy),aki_adm=as.factor(aki_adm),pre_rrt=as.factor(pre_rrt))

```


```{r}
df_1$ethnicity[df_1$ethnicity ==""| 
             df_1$ethnicity == "Native American"|
             df_1$ethnicity == "Other/Unknow"] <- "Other/Unknown"

df_1$surgery<-NA
df_1$surgery[df_1$hospitalAdmitSource =="Operating Room"| 
             df_1$hospitalAdmitSource == "PACU"] <- 1
df_1$surgery[df_1$hospitalAdmitSource =="Acute Care/Floor"| 
             df_1$hospitalAdmitSource == "Chest Pain Center"|
             df_1$hospitalAdmitSource == "Direct Admit"|
               df_1$hospitalAdmitSource == "Emergency Department"|
               df_1$hospitalAdmitSource == "Floor"|
               df_1$hospitalAdmitSource == "ICU"|
               df_1$hospitalAdmitSource == "ICU to SDU"|
               df_1$hospitalAdmitSource == "Observation"|
               df_1$hospitalAdmitSource == "Other"|
               df_1$hospitalAdmitSource == "Other Hospital"|
               df_1$hospitalAdmitSource == "Other ICU"|
               df_1$hospitalAdmitSource == "Recovery Room"|
               df_1$hospitalAdmitSource == "Step-Down Unit (SDU)"] <- 0
summary(df_1$hospitalAdmitSource)
summary(df_1$surgery)

#Categorizing Na, Cl, and SID
df_2<-df_1%>%mutate(cl_adm_cat=case_when(.$chloride_first<98 ~ 1,
                              .$chloride_first>=98 & .$chloride_first<=110 ~ 0,
                              .$chloride_first>110 ~ 2,
                                TRUE ~ as.numeric(NA)),
         na_adm_cat=case_when(.$sodium_first>145 ~ 2,
                                .$sodium_first<=145  & .$sodium_first>=135 ~ 0,
                                .$sodium_first<135 ~ 1,
                                TRUE ~ as.numeric(NA)),
         sid_adm_cat=case_when(.$SID_first>37 ~ 2,
                                .$SID_first<=37  & .$SID_first>=31 ~ 0,
                                .$SID_first<31 ~ 1,
                                TRUE ~ as.numeric(NA)),
         # detailed category
         cl_adm_catd=case_when(.$chloride_first>116 ~ 13,
                                   .$chloride_first>114&.$chloride_first<=116 ~ 12,
                                   .$chloride_first>112&.$chloride_first<=114 ~ 11,
                                   .$chloride_first>110&.$chloride_first<=112 ~ 10,
                                   .$chloride_first>108&.$chloride_first<=110 ~ 9,
                                   .$chloride_first>106&.$chloride_first<=108 ~ 8,
                                   .$chloride_first>104&.$chloride_first<=106 ~ 7,
                                   .$chloride_first>102&.$chloride_first<=104 ~ 6,
                                   .$chloride_first>100&.$chloride_first<=102 ~ 5,
                                   .$chloride_first>98&.$chloride_first<=100~ 4,
                                   .$chloride_first>96&.$chloride_first<=98 ~ 3,
                                   .$chloride_first>94&.$chloride_first<=96 ~ 2,
                                   .$chloride_first<=94~1,
                                   TRUE ~ as.numeric(NA)),
         na_adm_catd=case_when(.$sodium_first>146 ~ 11,
                                   .$sodium_first>144&.$sodium_first<=146 ~ 10,
                                   .$sodium_first>142&.$sodium_first<=144 ~ 9,
                                   .$sodium_first>140&.$sodium_first<=142 ~ 8,
                                   .$sodium_first>138&.$sodium_first<=140 ~ 7,
                                   .$sodium_first>136&.$sodium_first<=138 ~ 6,
                                   .$sodium_first>134&.$sodium_first<=136 ~ 5,
                                   .$sodium_first>132&.$sodium_first<=134 ~ 4,
                                   .$sodium_first>130&.$sodium_first<=132 ~ 3,
                                   .$sodium_first<=130~1,
                                   TRUE ~ as.numeric(NA)),
         sid_adm_catd=case_when(.$SID_first>42 ~ 10,
                                .$SID_first>40  & .$SID_first<=42 ~ 9,
                                .$SID_first>38  & .$SID_first<=40 ~ 8,
                                .$SID_first>36  & .$SID_first<=38 ~ 7,
                                .$SID_first>34  & .$SID_first<=36 ~ 6,
                                .$SID_first>32  & .$SID_first<=34 ~ 5,
                                .$SID_first>30  & .$SID_first<=32 ~ 4,
                                .$SID_first>28  & .$SID_first<=30 ~ 3,
                                .$SID_first>26  & .$SID_first<=28 ~ 2,
                                .$SID_first<=26 ~ 1,
                                TRUE ~ as.numeric(NA))
         )%>%mutate(gender=case_when(.$gender=="Female"~0,
                          .$gender=="Male"~1,
                          TRUE~as.numeric(NA)))%>%
   mutate_at(c("gender","pasthistory_chf","pasthistory_pvd","pasthistory_hypertension","pasthistory_copd","pasthistory_diabetes","cl_adm_cat","na_adm_cat","sid_adm_cat","cl_adm_catd","na_adm_catd","sid_adm_catd","ethnicity","surgery"),as.factor)
```


```{r}
#Exclude rrt before admision
df_3<-df_2%>%
  mutate(aki=case_when(.$AKIstage==0~0,.$AKIstage%in%c(1,2,3)~1))%>%
  mutate(pre_rrt=case_when(.$rrt_bin==1&first_rrtoffset<=0~1,
                           .$rrt_bin==0|first_rrtoffset>1~0, 
                          TRUE~as.numeric(NA)))%>%filter(pre_rrt==0|is.na(pre_rrt)) # There is not rrt_bin=0, so NA means no rrt?

nrow(df_2)-nrow(df_3)

# Exclude cre>=3 at admission
df_4<-df_3%>% filter(creat_first<=3|is.na(creat_first))

#Exclude aki at admission
#df_1<-df_1%>% mutate(aki_adm=case_when(.$creat_adm>=.$creat1+0.3|.$creat_adm>=.$creat1*1.5~1,
 #                                .$creat_adm<.$creat1+0.3&.$creat_adm<.$creat1*1.5~0,
  #                                     TRUE~as.numeric(NA)))%>%filter(aki_adm==1|is.na(aki_adm))

nrow(df_3)-nrow(df_4)

#Exclude surgical patients
df_me<-df_4%>%filter(surgery==0)

nrow(df_4)-nrow(df_me) # Number of surgical patients 

# Excluded patients
nrow(sid_aki_study)-nrow(df_me) 

```


# Imputation on selected variables
## Mean Imputation

```{r}
# Plan
## Complete case for predictor
## IPW for outcome
## Single, sinple imputation for continuous covariates
## Indicator variable for discrete covariates

# Select variables
df_me_selected<-df_me%>%dplyr::select(-creat_adm_offset,-first_rrtoffset,-peakcreat48h_offset,-peakcreat7d_offset,-peakcreat7d_to_discharge_offsetgap,-first_dialysis_in_the_icu_first7days_offset,-hospitalAdmitSource,-pre_rrt,-creat_adm)

df_me_selected_imputed<-df_me_selected

for(i in 1:ncol(df_me_selected_imputed)){
  df_me_selected_imputed[is.na(df_me_selected_imputed[,i]), i] <- mean(df_me_selected_imputed[,i], na.rm = TRUE)
}
```
### Studying variables missing values

```{r}
missing_values<-as.data.frame(sapply(df_me_selected, function(x) round(sum(is.na(x))/length(x),2)))
names(missing_values)<-'percentage_missing'

toomany_missing<-rownames(subset(missing_values,missing_values$percentage_missing>0.70))
toomany_missing
```

### Final replacement
We are not imputing values listed above, so we replace the imputed columns with the orginal ones
```{r}
keep.origninal<-dplyr::select(df_me_selected,c(patientunitstayid,
# MNAR  

# too many missing values

# Predictors and outcomes: No? impute them?
sodium_first,chloride_first,SID_first,cl_adm_cat,na_adm_cat,sid_adm_cat,cl_adm_catd,na_adm_catd,sid_adm_catd,
peakcreat48h,peakcreat7d,AKIstage,dialysis_in_the_icu_first7days_bin,unitlos_days ,hosplos_days ,ventdays_days,hospitalmortality,icumortality,
# Categorical variables should be dealed with using indicator variable, not imputed.
aki,AKIstage,surgery,gender,ethnicity,pasthistory_chf,pasthistory_pvd,pasthistory_hypertension,pasthistory_copd,pasthistory_diabetes,pasthistory_cld,pasthistory_malignancy
)
)

# Final replacement
df_me_selected_imputed[,names(keep.origninal)] <- keep.origninal

# Complete case for predictors
df_me_selected_imputed<-df_me_selected_imputed%>%filter(!is.na(sodium_first),!is.na(chloride_first),!is.na(SID_first))%>%droplevels()
```

# Count data
```{r}
colnames(df_me_selected_imputed)
df_me_selected_imputed%>%ggplot()+
  geom_histogram(aes(sid_adm_catd),stat="count")+scale_x_discrete(labels=c("-26","26-28","28-30","30-32","32-34","34-36","36-38","38-40","40-42","42-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=10))

df_me_selected_imputed%>%ggplot()+
  geom_histogram(aes(na_adm_catd),stat="count")+scale_x_discrete(labels=c("-130","130-132","132-134","134-136","136-138","138-140","140-142","142-144","144-146","146-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=10))

df_me_selected_imputed%>%ggplot()+
  geom_histogram(aes(cl_adm_catd),stat="count")+scale_x_discrete(labels=c("-94","94-96","96-98","98-100","100-102","102-104","104-106","106-108","108-110","110-112","112-114","114-116","116-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=10))
```



```{r}
# Patients' characteristics
df_me_selected_comp<-df_me_selected%>%filter(!is.na(sodium_first),!is.na(chloride_first),!is.na(SID_first))%>%mutate(aki=as.factor(aki))%>%mutate(icumortality=as.factor(icumortality),hospitalmortality=as.factor(hospitalmortality))%>%droplevels()
nrow(df_me_selected)-nrow(df_me_selected_comp)
# Sample size
nrow(df_me_selected_comp)
summary(df_me_selected_comp)
describe(df_me_selected_comp$age_fixed)
describe(df_me_selected_comp$weight)
summary(df_me_selected_comp$creat_first)

#SID, Na, and Cl
summary(df_me_selected_comp$sid_adm_cat)
prop.table(summary(df_me_selected_comp$sid_adm_cat))
summary(df_me_selected_comp$na_adm_cat)
prop.table(summary(df_me_selected_comp$na_adm_cat))
summary(df_me_selected_comp$cl_adm_cat)
prop.table(summary(df_me_selected_comp$cl_adm_cat))

# AKI
summary(df_me_selected_comp$aki)
prop.table(summary(df_me_selected_comp$aki))
df_me_selected_comp%>% mutate(AKIstage=as.factor(AKIstage))%>%dplyr::select(AKIstage)%>%summary()

# Mortality
summary(df_me_selected_comp$icumortality)
prop.table(summary(df_me_selected_comp$icumortality))

summary(df_me_selected_comp$hospitalmortality)
prop.table(summary(df_me_selected_comp$hospitalmortality))

# Table one: base
table<-CreateTableOne(vars=c("age_fixed","gender","height","weight","ethnicity","pasthistory_chf","pasthistory_pvd","pasthistory_hypertension","pasthistory_copd","pasthistory_diabetes","pasthistory_malignancy","creat_first"),strata="sid_adm_cat",data=df_me_selected_comp)

try <- print(table, exact = "ascites", quote = FALSE)
write.csv(try, file = "table.csv")

# Table one: Characteristics
table<-CreateTableOne(vars=c("SID_first","chloride_first","sodium_first","albumin_first","bilirubin_first","BUN_first","wbc_first","platelets_first","hematocrit_first","hemoglobin_first","glucose_first","potassium_adm","ua_adm","heartrate_adm","systolic_adm","temperature_adm","bicarbonate_adm"),strata="sid_adm_cat",data=df_me_selected_comp)
try <- print(table, exact = "ascites", quote = FALSE)
write.csv(try, file = "table.csv")

colnames(df_me_selected_comp)

# Table one: Outcome
table<-CreateTableOne(vars=c("aki","AKIstage","icumortality",'hospitalmortality','unitlos_days' ,'hosplos_days'),strata="sid_adm_cat",data=df_me_selected_comp)
try <- print(table, exact = "ascites", quote = FALSE)
write.csv(try, file = "table.csv")

```


```{r}
write.csv(df_me_selected_imputed,"df_me_selected_imputed.csv")
write.csv(df_me_selected_comp,"df_me_selected_comp.csv")
```

# Univariable analyses
## AKI
```{r}
df_me_selected_imputed<-df_me_selected_imputed%>%mutate(aki=as.numeric(as.character(aki)))
#SID
# Upper figure
p_1<-df_me_selected_imputed%>%filter(!is.na(sid_adm_catd))%>%group_by(sid_adm_catd)%>%summarize(aki_incidence=sum(aki,na.rm = T)/n()*100)%>%ggplot()+
  geom_bar(aes(x=sid_adm_catd,y=aki_incidence),width=0.8,stat="identity")+
  scale_x_discrete(labels=c("-26","26-28","28-30","30-32","32-34","34-36","36-38","38-40","40-42","42-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=12),axis.text.y = element_text(size=12))+
  xlab(expression(paste(mSID," (mEq/L)")))+
  ylab(label="AKI incidence (%)")

# Table
l_1<-df_me_selected_imputed%>%filter(!is.na(sid_adm_catd))%>%group_by(sid_adm_catd)%>%summarize(n=n(),death=sum(aki,na.rm = T))
c1<-data.frame(SID=c("-26","26-28","28-30","30-32","32-34","34-36","36-38","38-40","40-42","42-"))
l_1<-cbind(c1,l_1)%>%dplyr::select(-sid_adm_catd)
tab = t(l_1)
tab = tableGrob(tab, rows=NULL,theme=ttheme_default(base_size = 12))
tab$widths <- unit(rep(1/ncol(tab), ncol(tab)), "npc")

# Legend of table
le <- structure(list(a = "SID (mEq/L)", b="n", c ="AKI"), .Names = c("SID","N", "AKI"), class = "data.frame", row.names = c(NA, -1L))
le_tab = t(le)
le_tab = tableGrob(le_tab, rows=NULL,theme=ttheme_default(base_size = 12))
le_tab$widths <- unit(rep(1/ncol(le_tab), ncol(le_tab)), "npc")

# Combine
grid.arrange(arrangeGrob(nullGrob(), p_1 + guides(fill=FALSE),widths=c(1,8)), 
         arrangeGrob(arrangeGrob(nullGrob(),le_tab,heights=c(1,0),widths=c(1,1)), tab, nullGrob(), widths=c(10,45,1)), heights=c(8,1)) 
```


```{r}
#Cl
# Upper figure
p_1<-df_me_selected_imputed%>%filter(!is.na(cl_adm_catd))%>%group_by(cl_adm_catd)%>%summarize(aki_incidence=sum(aki,na.rm =T)/n()*100)%>%ggplot()+geom_bar(aes(x=cl_adm_catd,y=aki_incidence),width=0.8,stat="identity")+scale_x_discrete(labels=c("-94","94-96","96-98","98-100","100-102","102-104","104-106","106-108","108-110","110-112","112-114","114-116","116-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=12),axis.text.y = element_text(size=12))+xlab(expression(paste("[Cl"^"-","] (mEq/L)")))+ ylab(label="AKI incidence (%)")


# Table
l_1<-df_me_selected_imputed%>%filter(!is.na(cl_adm_catd))%>%group_by(cl_adm_catd)%>%summarize(n=n(),death=sum(aki,na.rm = T))
c1<-data.frame(time=c("-94","94-96","96-98","98-100","100-102","102-104","104-106","106-108","108-110","110-112","112-114","114-116","116-"))
l_1<-cbind(c1,l_1)%>%dplyr::select(-cl_adm_catd)
tab = t(l_1)
tab = tableGrob(tab, rows=NULL, theme=ttheme_default(base_size = 12))
tab$widths <- unit(rep(1/ncol(tab), ncol(tab)), "npc")

# Legend of table
le <- structure(list(a = "[Cl](mEq/L)", b ="n", c="AKI"), .Names = c("SID","N", "AKI"), class = "data.frame", row.names = c(NA, -1L))
le_tab = t(le)
le_tab = tableGrob(le_tab, rows=NULL,theme=ttheme_default(base_size = 12))
le_tab$widths <- unit(rep(1/ncol(le_tab), ncol(le_tab)), "npc")

# Combine
grid.arrange(arrangeGrob(nullGrob(), p_1 + guides(fill=FALSE),widths=c(1,8)), 
         arrangeGrob(arrangeGrob(nullGrob(),le_tab,heights=c(1,0),widths=c(1,1)), tab, nullGrob(), widths=c(9,45,1)), heights=c(8,1)) 



#Na
#df_me_selected_imputed%>%filter(!is.na(na_adm_catd))%>%group_by(na_adm_catd)%>%summarize(aki_incidence=sum(aki,na.rm = T)/n()*100)%>%ggplot()+geom_bar(aes(x=na_adm_catd,y=aki_incidence),width=0.8,stat="identity")+scale_x_discrete(labels=c("-130","130-132","132-134","134-136","136-138","138-140","140-142","142-144","144-146","146-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=10))+xlab(expression(paste("Na"^"+"," (mEq/L)")))+ ylab(label="AKI incidence (%)")
```


# Multivariable regression
## AKI
### Whole data without IPW
### Without bicarbonate, because if included, interpretation is difficult
```{r}
# SID and Cl without IPW
model<-glm(aki~C(sid_adm_catd,base=5)+C(cl_adm_catd,base=7)+
age_fixed+gender+weight+height+ethnicity+pasthistory_chf+pasthistory_pvd+pasthistory_hypertension+pasthistory_copd+pasthistory_diabetes+pasthistory_malignancy+heartrate_adm+systolic_adm+temperature_adm+wbc_first+platelets_first+hematocrit_first+hemoglobin_first+albumin_first+bilirubin_first+BUN_first+glucose_first+creat_first+potassium_adm+ua_adm,data=df_me_selected_imputed,family="binomial")
summary(model)

try<-as.data.frame(summary(model)%>%coef())%>%round(3)
write.csv(try, file = "table.csv")
table<-exp(cbind(OR=coef(model), confint.default(model)))%>%round(2)
try<-print(table, exact = "ascites", quote = FALSE)
write.csv(try, file = "table.csv")

as.data.frame(summary(model)%>%coef())%>%round(6)
exp(0.011737*10)
coef(model)[25]
exp(cbind(coef(model)[23],confint.default(model)[23]))
# Age
coef(model)[23]
exp(cbind(coef(model)*10,confint.default(model)*10))[23,]
# Weight
coef(model)[25]
exp(cbind(coef(model)*10,confint.default(model)*10))[25,]
# Height
coef(model)[26]
exp(cbind(coef(model)*10,confint.default(model)*10))[26,]
# HR
coef(model)[37]
exp(cbind(coef(model)*10,confint.default(model)*10))[37,]
# SBP
coef(model)[38]
exp(cbind(coef(model)*10,confint.default(model)*10))[38,]
# BUN
coef(model)[46]
exp(cbind(coef(model)*10,confint.default(model)*10))[46,]
# Glu
coef(model)[47]
exp(cbind(coef(model)*20,confint.default(model)*20))[47,]

# Plt
coef(model)[41]
exp(cbind(coef(model)*41,confint.default(model)*20))[41,]
```


### Subgroup without IPW
```{r}
df_me_selected_imputed_sub<-df_me_selected_imputed%>%mutate(sid_adm_catd=as.numeric(as.character(sid_adm_catd)),cl_adm_catd=as.numeric(as.character(cl_adm_catd)))
df_sid_l_cl_l<-df_me_selected_imputed_sub%>%filter(SID_first<=34&chloride_first<=106)
df_sid_l_cl_h<-df_me_selected_imputed_sub%>%filter(SID_first<=34&chloride_first>106)
df_sid_h_cl_l<-df_me_selected_imputed_sub%>%filter(SID_first>34&chloride_first<=106)
df_sid_h_cl_h<-df_me_selected_imputed_sub%>%filter(SID_first>34&chloride_first>106)
summary(df_me_selected_imputed$cl_adm_catd)
```

```{r}
summary(df_sid_l_cl_l)

# Low SID and Low Cl
model<-glm(aki~SID_first+chloride_first+
age_fixed+gender+weight+height+albumin_first+bilirubin_first+BUN_first+wbc_first+platelets_first+hematocrit_first+hemoglobin_first+glucose_first+ethnicity+pasthistory_chf+pasthistory_pvd+pasthistory_hypertension+pasthistory_copd+pasthistory_diabetes+pasthistory_malignancy+creat_first+potassium_adm+ua_adm+heartrate_adm+systolic_adm+temperature_adm,data=df_sid_l_cl_l,family="binomial")
summary(model)
exp(cbind(OR=coef(model), confint.default(model)))%>%round(2)

# Low SID and High Cl
model<-glm(aki~SID_first+chloride_first+
age_fixed+gender+weight+height+albumin_first+bilirubin_first+BUN_first+wbc_first+platelets_first+hematocrit_first+hemoglobin_first+glucose_first+ethnicity+pasthistory_chf+pasthistory_pvd+pasthistory_hypertension+pasthistory_copd+pasthistory_diabetes+pasthistory_malignancy+creat_first+potassium_adm+ua_adm+heartrate_adm+systolic_adm+temperature_adm,data=df_sid_l_cl_h,family="binomial")
summary(model)
exp(cbind(OR=coef(model), confint.default(model)))%>%round(2)

# High SID and Low Cl
model<-glm(aki~SID_first+chloride_first+
age_fixed+gender+weight+height+albumin_first+bilirubin_first+BUN_first+wbc_first+platelets_first+hematocrit_first+hemoglobin_first+glucose_first+ethnicity+pasthistory_chf+pasthistory_pvd+pasthistory_hypertension+pasthistory_copd+pasthistory_diabetes+pasthistory_malignancy+creat_first+potassium_adm+ua_adm+heartrate_adm+systolic_adm+temperature_adm,data=df_sid_h_cl_l,family="binomial")
summary(model)
exp(cbind(OR=coef(model), confint.default(model)))%>%round(2)
# High SID and High Cl
model<-glm(aki~SID_first+chloride_first+
age_fixed+gender+weight+height+albumin_first+bilirubin_first+BUN_first+wbc_first+platelets_first+hematocrit_first+hemoglobin_first+glucose_first+ethnicity+pasthistory_chf+pasthistory_pvd+pasthistory_hypertension+pasthistory_copd+pasthistory_diabetes+pasthistory_malignancy+creat_first+potassium_adm+ua_adm+heartrate_adm+systolic_adm+temperature_adm,data=df_sid_h_cl_h,family="binomial")
summary(model)
exp(cbind(OR=coef(model), confint.default(model)))%>%round(2)
```


### IPW
```{r}
# SID and Cl for new categories without bicarbonate, with IPW
# "aki" is outcome with missing value
# Both "sid_adm_catd" and "cl_adm_catd" are categorical predictors of interest

# Create censoring variable
df_me_selected_imputed_ipw<-df_me_selected_imputed
df_me_selected_imputed_ipw$cens<-as.numeric(is.na(df_me_selected_imputed_ipw$aki))

# Calculate denominotor of IPW for selection bias
denom.cens <- glm(cens ~C(sid_adm_catd,base=5)+C(cl_adm_catd,base=7)+
age_fixed+weight+height+albumin_first+bilirubin_first+BUN_first+wbc_first+platelets_first+hematocrit_first+hemoglobin_first+glucose_first+creat_first+potassium_adm+ua_adm+heartrate_adm+systolic_adm+temperature_adm,family="binomial", data = df_me_selected_imputed_ipw)

pd.cens <- 1-predict(denom.cens, df_me_selected_imputed_ipw, type = "response")

# Calculate numerator of IPW for selection bias
numer.cens <- glm(cens~C(sid_adm_catd,base=5)+C(cl_adm_catd,base=7), family = binomial(), data = df_me_selected_imputed_ipw)
pn.cens <- 1-predict(numer.cens, df_me_selected_imputed_ipw, type = "response")
df_me_selected_imputed_ipw$sw.c <- pn.cens/pd.cens


# Final model
# Maybe,to use "geeglm", there should be no missing value; stage 2 regression is for complete case subgroup.
df_me_selected_imputed_ipw<-df_me_selected_imputed_ipw%>%filter(!is.na(gender)&!is.na(ethnicity)&!is.na(pasthistory_chf)&!is.na(pasthistory_pvd)&!is.na(pasthistory_hypertension)&!is.na(pasthistory_copd)&!is.na(pasthistory_diabetes)&!is.na(pasthistory_malignancy))

model <- geeglm(aki ~ C(sid_adm_catd,base=5)+C(cl_adm_catd,base=7)+
age_fixed+gender+weight+height+albumin_first+bilirubin_first+BUN_first+wbc_first+platelets_first+hematocrit_first+hemoglobin_first+glucose_first+ethnicity+pasthistory_chf+pasthistory_pvd+pasthistory_hypertension+pasthistory_copd+pasthistory_diabetes+pasthistory_malignancy+creat_first+potassium_adm+ua_adm+heartrate_adm+systolic_adm+temperature_adm, data=df_me_selected_imputed_ipw, weights=sw.c, id=patientunitstayid,corstr="independence",family= binomial(link = "logit"))

summary(model)
try<-as.data.frame(summary(model)%>%coef())%>%round(3)
write.csv(try, file = "table.csv")
OR<-exp(coef(model))
SE <- coef(summary(model))[,2]
LCI<-exp(coef(model)-qnorm(0.975)*SE)
UCI<-exp(coef(model)+qnorm(0.975)*SE)
try<-cbind(OR, LCI, UCI)%>%round(2)
write.csv(try, file = "table.csv")
```





## ICU Morality
```{r}
df_me_selected_imputed<-df_me_selected_imputed%>%mutate(icumortality=as.numeric(as.character(icumortality)))
#SID
df_me_selected_imputed%>%filter(!is.na(sid_adm_catd))%>%group_by(sid_adm_catd)%>%summarize(icumortality_incidence=sum(icumortality,na.rm = T)/n()*100)%>%ggplot()+
  geom_bar(aes(x=sid_adm_catd,y=icumortality_incidence),width=0.8,stat="identity")+
  scale_x_discrete(labels=c("-26","26-28","28-30","30-32","32-34","34-36","36-38","38-40","40-42","42-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=10))+
  xlab(expression(paste(mSID," (mEq/L)")))+
  ylab(label="ICU mortality (%)")


#Cl
df_me_selected_imputed%>%filter(!is.na(cl_adm_catd))%>%group_by(cl_adm_catd)%>%summarize(icumortality_incidence=sum(icumortality,na.rm = T)/n()*100)%>%ggplot()+geom_bar(aes(x=cl_adm_catd,y=icumortality_incidence),width=0.8,stat="identity")+scale_x_discrete(labels=c("-94","94-96","96-98","98-100","100-102","102-104","104-106","106-108","108-110","110-112","112-114","114-116","116-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=10))+xlab(expression(paste("Cl"^"-"," (mEq/L)")))+ ylab(label="ICU mortality (%)")

#Na
#df_me_selected_imputed%>%filter(!is.na(na_24_mean_catd))%>%group_by(na_24_mean_catd)%>%summarize(icumortality_incidence=sum(icumortality,na.rm = T)/n()*100)%>%ggplot()+geom_bar(aes(x=na_24_mean_catd,y=icumortality_incidence),width=0.8,stat="identity")+scale_x_discrete(labels=c("-130","130-132","132-134","134-136","136-138","138-140","140-142","142-144","144-146","146-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=10))+xlab(expression(paste("Na"^"+"," (mEq/L)")))+ ylab(label="ICU mortality (%)")
```


```{r}
# SID
# Upper figure
p_1<-df_me_selected_imputed%>%filter(!is.na(sid_adm_catd))%>%group_by(sid_adm_catd)%>%summarize(icumortality_incidence=sum(icumortality,na.rm = T)/n()*100)%>%ggplot()+
  geom_bar(aes(x=sid_adm_catd,y=icumortality_incidence),width=0.8,stat="identity")+
  scale_x_discrete(labels=c("-26","26-28","28-30","30-32","32-34","34-36","36-38","38-40","40-42","42-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=12),axis.text.y = element_text(size=12))+
  xlab(expression(paste(mSID," (mEq/L)")))+
  ylab(label="ICU mortality (%)")

# Table
l_1<-df_me_selected_imputed%>%filter(!is.na(sid_adm_catd))%>%group_by(sid_adm_catd)%>%summarize(n=n(),death=sum(icumortality,na.rm = T))
c1<-data.frame(SID=c("-26","26-28","28-30","30-32","32-34","34-36","36-38","38-40","40-42","42-"))
l_1<-cbind(c1,l_1)%>%dplyr::select(-sid_adm_catd)
tab = t(l_1)
tab = tableGrob(tab, rows=NULL,theme=ttheme_default(base_size = 12))
tab$widths <- unit(rep(1/ncol(tab), ncol(tab)), "npc")

# Legend of table
le <- structure(list(a = "SID (mEq/L)", b="n", c ="Death"), .Names = c("SID","N", "DEATH"), class = "data.frame", row.names = c(NA, -1L))
le_tab = t(le)
le_tab = tableGrob(le_tab, rows=NULL,theme=ttheme_default(base_size = 12))
le_tab$widths <- unit(rep(1/ncol(le_tab), ncol(le_tab)), "npc")

# Combine
grid.arrange(arrangeGrob(nullGrob(), p_1 + guides(fill=FALSE),widths=c(1,8)), 
         arrangeGrob(arrangeGrob(nullGrob(),le_tab,heights=c(1,0),widths=c(1,1)), tab, nullGrob(), widths=c(10,45,1)), heights=c(8,1)) 
```


```{r}
# Cl
# Upper figure
p_1<-df_me_selected_imputed%>%filter(!is.na(cl_adm_catd))%>%group_by(cl_adm_catd)%>%summarize(icumortality_incidence=sum(icumortality,na.rm = T)/n()*100)%>%ggplot()+geom_bar(aes(x=cl_adm_catd,y=icumortality_incidence),width=0.8,stat="identity")+scale_x_discrete(labels=c("-94","94-96","96-98","98-100","100-102","102-104","104-106","106-108","108-110","110-112","112-114","114-116","116-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=10),axis.text.y = element_text(size=12))+xlab(expression(paste("[Cl"^"-","] (mEq/L)")))+ ylab(label="ICU mortality (%)")

# Table
l_1<-df_me_selected_imputed%>%filter(!is.na(cl_adm_catd))%>%group_by(cl_adm_catd)%>%summarize(n=n(),death=sum(icumortality,na.rm = T))
c1<-data.frame(time=c("-94","94-96","96-98","98-100","100-102","102-104","104-106","106-108","108-110","110-112","112-114","114-116","116-"))
l_1<-cbind(c1,l_1)%>%dplyr::select(-cl_adm_catd)
tab = t(l_1)
tab = tableGrob(tab, rows=NULL, theme=ttheme_default(base_size = 12))
tab$widths <- unit(rep(1/ncol(tab), ncol(tab)), "npc")

# Legend of table
le <- structure(list(a = "[Cl] (mEq/L)", b ="n", c="Death"), .Names = c("SID","N", "DEATH"), class = "data.frame", row.names = c(NA, -1L))
le_tab = t(le)
le_tab = tableGrob(le_tab, rows=NULL,theme=ttheme_default(base_size = 12))
le_tab$widths <- unit(rep(1/ncol(le_tab), ncol(le_tab)), "npc")

# Combine
grid.arrange(arrangeGrob(nullGrob(), p_1 + guides(fill=FALSE),widths=c(1,8)), 
         arrangeGrob(arrangeGrob(nullGrob(),le_tab,heights=c(1,0),widths=c(1,1)), tab, nullGrob(), widths=c(9,45,1)), heights=c(8,1)) 

```




### Whole data without IPW for ICU mortaltiy
### Without bicarbonate, because if included, interpretation is difficult
```{r}
# SID and Cl without IPW
model<-glm(icumortality~C(sid_adm_catd,base=5)+C(cl_adm_catd,base=7)+
age_fixed+gender+weight+height+ethnicity+pasthistory_chf+pasthistory_pvd+pasthistory_hypertension+pasthistory_copd+pasthistory_diabetes+pasthistory_malignancy+heartrate_adm+systolic_adm+temperature_adm+wbc_first+platelets_first+hematocrit_first+hemoglobin_first+albumin_first+bilirubin_first+BUN_first+glucose_first+creat_first+potassium_adm+ua_adm,data=df_me_selected_imputed,family="binomial")
summary(model)

try<-as.data.frame(summary(model)%>%coef())%>%round(3)
write.csv(try, file = "table.csv")
table<-exp(cbind(OR=coef(model), confint.default(model)))%>%round(2)
try<-print(table, exact = "ascites", quote = FALSE)
write.csv(try, file = "table.csv")

# Age
coef(model)[23]
exp(cbind(coef(model)*10,confint.default(model)*10))[23,]
# Weight
coef(model)[25]
exp(cbind(coef(model)*10,confint.default(model)*10))[25,]
# Height
coef(model)[26]
exp(cbind(coef(model)*10,confint.default(model)*10))[26,]
# SBP
coef(model)[38]
exp(cbind(coef(model)*10,confint.default(model)*10))[38,]
# BUN
coef(model)[46]
exp(cbind(coef(model)*10,confint.default(model)*10))[46,]
# Glu
coef(model)[47]
exp(cbind(coef(model)*20,confint.default(model)*20))[47,]

# Plt
coef(model)[41]
exp(cbind(coef(model)*41,confint.default(model)*20))[41,]
```

### Subgroup without IPW
```{r}
summary(df_sid_l_cl_l)

# Low SID and Low Cl
model<-glm(icumortality~SID_first+chloride_first+
age_fixed+gender+weight+height+albumin_first+bilirubin_first+BUN_first+wbc_first+platelets_first+hematocrit_first+hemoglobin_first+glucose_first+ethnicity+pasthistory_chf+pasthistory_pvd+pasthistory_hypertension+pasthistory_copd+pasthistory_diabetes+pasthistory_malignancy+creat_first+potassium_adm+ua_adm+heartrate_adm+systolic_adm+temperature_adm,data=df_sid_l_cl_l,family="binomial")
summary(model)
exp(cbind(OR=coef(model), confint.default(model)))%>%round(2)

# Low SID and High Cl
model<-glm(icumortality~SID_first+chloride_first+
age_fixed+gender+weight+height+albumin_first+bilirubin_first+BUN_first+wbc_first+platelets_first+hematocrit_first+hemoglobin_first+glucose_first+ethnicity+pasthistory_chf+pasthistory_pvd+pasthistory_hypertension+pasthistory_copd+pasthistory_diabetes+pasthistory_malignancy+creat_first+potassium_adm+ua_adm+heartrate_adm+systolic_adm+temperature_adm,data=df_sid_l_cl_h,family="binomial")
summary(model)
exp(cbind(OR=coef(model), confint.default(model)))%>%round(2)

# High SID and Low Cl
model<-glm(icumortality~SID_first+chloride_first+
age_fixed+gender+weight+height+albumin_first+bilirubin_first+BUN_first+wbc_first+platelets_first+hematocrit_first+hemoglobin_first+glucose_first+ethnicity+pasthistory_chf+pasthistory_pvd+pasthistory_hypertension+pasthistory_copd+pasthistory_diabetes+pasthistory_malignancy+creat_first+potassium_adm+ua_adm+heartrate_adm+systolic_adm+temperature_adm,data=df_sid_h_cl_l,family="binomial")
summary(model)
exp(cbind(OR=coef(model), confint.default(model)))%>%round(2)
# High SID and High Cl
model<-glm(icumortality~SID_first+chloride_first+
age_fixed+gender+weight+height+albumin_first+bilirubin_first+BUN_first+wbc_first+platelets_first+hematocrit_first+hemoglobin_first+glucose_first+ethnicity+pasthistory_chf+pasthistory_pvd+pasthistory_hypertension+pasthistory_copd+pasthistory_diabetes+pasthistory_malignancy+creat_first+potassium_adm+ua_adm+heartrate_adm+systolic_adm+temperature_adm,data=df_sid_h_cl_h,family="binomial")
summary(model)
exp(cbind(OR=coef(model), confint.default(model)))%>%round(2)
```



## Hospital mortality

```{r}
# SID
# Upper figure
p_1<-df_me_selected_imputed%>%filter(!is.na(sid_adm_catd))%>%group_by(sid_adm_catd)%>%summarize(hospitalmortality_incidence=sum(hospitalmortality,na.rm = T)/n()*100)%>%ggplot()+
  geom_bar(aes(x=sid_adm_catd,y=hospitalmortality_incidence),width=0.8,stat="identity")+
  scale_x_discrete(labels=c("-26","26-28","28-30","30-32","32-34","34-36","36-38","38-40","40-42","42-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=10))+
  xlab(expression(paste(mSID," (mEq/L)")))+
  ylab(label="Hospital mortality (%)")

# Table
l_1<-df_me_selected_imputed%>%filter(!is.na(sid_adm_catd))%>%group_by(sid_adm_catd)%>%summarize(n=n(),death=sum(hospitalmortality,na.rm = T))
c1<-data.frame(SID=c("-26","26-28","28-30","30-32","32-34","34-36","36-38","38-40","40-42","42-"))
l_1<-cbind(c1,l_1)%>%dplyr::select(-sid_adm_catd)
tab = t(l_1)
tab = tableGrob(tab, rows=NULL,theme=ttheme_default(base_size = 10))
tab$widths <- unit(rep(1/ncol(tab), ncol(tab)), "npc")

# Legend of table
le <- structure(list(a = "SID (mEq/L)", b="n", c ="Death"), .Names = c("SID","N", "DEATH"), class = "data.frame", row.names = c(NA, -1L))
le_tab = t(le)
le_tab = tableGrob(le_tab, rows=NULL,theme=ttheme_default(base_size = 10))
le_tab$widths <- unit(rep(1/ncol(le_tab), ncol(le_tab)), "npc")

# Combine
grid.arrange(arrangeGrob(nullGrob(), p_1 + guides(fill=FALSE),widths=c(1,8)), 
         arrangeGrob(arrangeGrob(nullGrob(),le_tab,heights=c(1,0),widths=c(1,1)), tab, nullGrob(), widths=c(10,45,1)), heights=c(8,1)) 

```

```{r}
# Cl
# Upper figure
p_1<-df_me_selected_imputed%>%filter(!is.na(cl_adm_catd))%>%group_by(cl_adm_catd)%>%summarize(hospitalmortality_incidence=sum(hospitalmortality,na.rm = T)/n()*100)%>%ggplot()+geom_bar(aes(x=cl_adm_catd,y=icumortality_incidence),width=0.8,stat="identity")+scale_x_discrete(labels=c("-94","94-96","96-98","98-100","100-102","102-104","104-106","106-108","108-110","110-112","112-114","114-116","116-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=10))+xlab(expression(paste("Cl"^"-"," (mEq/L)")))+ ylab(label="Hospital mortality (%)")

# Table
l_1<-df_me_selected_imputed%>%filter(!is.na(cl_adm_catd))%>%group_by(cl_adm_catd)%>%summarize(n=n(),death=sum(hospitalmortality,na.rm = T))
c1<-data.frame(time=c("-94","94-96","96-98","98-100","100-102","102-104","104-106","106-108","108-110","110-112","112-114","114-116","116-"))
l_1<-cbind(c1,l_1)%>%dplyr::select(-cl_adm_catd)
tab = t(l_1)
tab = tableGrob(tab, rows=NULL, theme=ttheme_default(base_size = 8))
tab$widths <- unit(rep(1/ncol(tab), ncol(tab)), "npc")

# Legend of table
le <- structure(list(a = "Cl (mEq/L)", b ="n", c="Death"), .Names = c("SID","N", "DEATH"), class = "data.frame", row.names = c(NA, -1L))
le_tab = t(le)
le_tab = tableGrob(le_tab, rows=NULL,theme=ttheme_default(base_size = 8))
le_tab$widths <- unit(rep(1/ncol(le_tab), ncol(le_tab)), "npc")

# Combine
grid.arrange(arrangeGrob(nullGrob(), p_1 + guides(fill=FALSE),widths=c(1,8)), 
         arrangeGrob(arrangeGrob(nullGrob(),le_tab,heights=c(1,0),widths=c(1,1)), tab, nullGrob(), widths=c(9,45,1)), heights=c(8,1)) 

```


```{r}
df_me_selected_imputed<-df_me_selected_imputed%>%mutate(hospitalmortality=as.numeric(as.character(hospitalmortality)))
#SID
df_me_selected_imputed%>%filter(!is.na(hospitalmortality),!is.na(sid_adm_catd))%>%group_by(sid_adm_catd)%>%summarize(hospitalmortality_incidence=sum(hospitalmortality)/n()*100)%>%ggplot()+
  geom_bar(aes(x=sid_adm_catd,y=hospitalmortality_incidence),width=0.8,stat="identity")+
  scale_x_discrete(labels=c("-26","26-28","28-30","30-32","32-34","34-36","36-38","38-40","40-42","42-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=10))+
  xlab(expression(paste(mSID," (mEq/L)")))+
  ylab(label="Hospital mortality (%)")


#Cl
df_me_selected_imputed%>%filter(!is.na(hospitalmortality),!is.na(cl_adm_catd))%>%group_by(cl_adm_catd)%>%summarize(hospitalmortality_incidence=sum(hospitalmortality)/n()*100)%>%ggplot()+geom_bar(aes(x=cl_adm_catd,y=hospitalmortality_incidence),width=0.8,stat="identity")+scale_x_discrete(labels=c("-94","94-96","96-98","98-100","100-102","102-104","104-106","106-108","108-110","110-112","112-114","114-116","116-"))+ theme(axis.text.x = element_text(angle = 45, hjust = 1, size=10))+xlab(expression(paste("Cl"^"-"," (mEq/L)")))+ ylab(label="Hospital mortality (%)")
```

### Regression model
### Whole data without IPW for hospital mortality
### Without bicarbonate, because if included, interpretation is difficult
```{r}
# SID and Cl without IPW
model<-glm(hospitalmortality~C(sid_adm_catd,base=5)+C(cl_adm_catd,base=7)+
age_fixed+gender+weight+height+ethnicity+pasthistory_chf+pasthistory_pvd+pasthistory_hypertension+pasthistory_copd+pasthistory_diabetes+pasthistory_malignancy+heartrate_adm+systolic_adm+temperature_adm+wbc_first+platelets_first+hematocrit_first+hemoglobin_first+albumin_first+bilirubin_first+BUN_first+glucose_first+creat_first+potassium_adm+ua_adm,data=df_me_selected_imputed,family="binomial")
summary(model)

try<-as.data.frame(summary(model)%>%coef())%>%round(3)
write.csv(try, file = "table.csv")
table<-exp(cbind(OR=coef(model), confint.default(model)))%>%round(2)
try<-print(table, exact = "ascites", quote = FALSE)
write.csv(try, file = "table.csv")

```




##############
# Additional analyses
```{r}
colnames(df_me_selected_imputed)
ggplot(df_me_selected_imputed,aes(SID_first,bicarbonate_adm))+
  geom_point(alpha=0.3)+
  xlab(expression(paste(mSID," (mEq/L)")))+
  ylab(label="Serum bicarbonate (mEq/L)")

cor.test(df_me_selected_imputed$SID_first, df_me_selected_imputed$bicarbonate_adm, method="pearson", conf.level=0.95)

ggplot(df_me_selected_imputed,aes(SID_first,chloride_first))+
  geom_point(alpha=0.5)+
  ylab(expression(paste("[Cl]"^"-"," (mEq/L)")))+
  xlab(expression(paste(mSID," (mEq/L)")))

cor.test(df_me_selected_imputed$SID_first, df_me_selected_imputed$chloride_first, method="pearson", conf.level=0.95)


write.csv(df_me_selected_imputed,"test.csv")
```










